Common sense approaches to sharing tabular data alongside publication

### 2021-09-20 new-paper.Rmd
push companion files and make sure the paper compiles to word

---
title: "Common sense approaches to sharing tabular data alongside publication"
author: Nicholas J. Tierney\* (1, 2, 3) & Karthik Ram\* (4)
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
citation_package: 'natbib'
csl: nature-communications.csl
header-includes:
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

@@

Tools that can be used to prepare datasets and metadata (*the infrastructure*) and minimal sets of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015]. Unlike code, datasets are far more diverse in terms of heterogeneity, size, and format. This makes them particularly challenging to standardize or easily "install" where the code is running. There have been some efforts to improve this process, such as gogetdata, the Dat Project, Frictionless Data [@Fowler_Barratt_Walsh_2018], and the formation of a private company, Quilt.

@@

Journals provide data papers in a formal publication explaining data that is archived (e.g. Ecology, Scientific Data). This is a familiar avenue for researchers to receive credit and citation. Writing a data paper can be very useful for researchers, but, data papers lack guidelines on how to structure data for effective reuse and sharing. We recommend partnering with best practice data repositories that emphasize curation, like Dryad. This way the data will have a long-term archiving and will also have professional curation to ensure that the published data are usable and have enough metadata to be understood. While writing a data paper can be very useful for researchers, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing. Data journals should routinely be partnering with best practice repositories to best support the underlying data.

@@

One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits, such as preservation, indexing,  discoverability, and most importantly, citation with a persistent identifier. Authors should add a CITATION file when possible but the real burden lies with publishers to ensure that these citations are marked up appropriately to allow for the building of open data metrics [@lowenberg_2021, @lowenberg_daniella_2021b].

When citing data, you should cite the preferred CITATION file in the first instance, else, a data publication, (for example in _Scientific Data, Ecology_), and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).

>>>>

---
title: "Common sense approaches to sharing tabular data alongside publication"
author: Nicholas J. Tierney\* (1, 2, 3) & Karthik Ram\* (4)
output:
  bookdown::word_document2:
  bookdown::pdf_document2:
    keep_tex: true
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
citation_package: 'natbib'
csl: nature-communications.csl
header-includes:
  - \setstretch{1.45}
  - \usepackage[utf8]{inputenc}
---

@@

Tools that can be used to prepare datasets and metadata (*the infrastructure*) and minimal sets of practices (*guidelines*) for reuse are sparse [@Arslan2019; @Roche2015]. Unlike code, datasets are far more diverse in terms of heterogeneity, size, and format. This makes them particularly challenging to standardize or easily "install" where the code is running. There have been some efforts to improve this process, such as gogetdata, the Dat Project, and the formation of a private company, Quilt.

@@

Journals provide data papers in a formal publication explaining data that is archived. This is a familiar avenue for researchers to receive credit and citation. Writing a data paper can be very useful for researchers, but, data papers lack guidelines on how to structure data for effective reuse and sharing. Partner with Dryad to link archived data with research papers. Dryad will also provide professionals to curate your data, thereby improving usability. While writing a data paper can be very useful for researchers, journals that publish data papers lack guidelines on how to structure data for effective reuse and sharing.

@@

One critical advantage of depositing data in a trusted repository (Section \@ref(avenues)) is that it comes with many benefits, such as preservation, indexing,  discoverability, and most importantly, citation with a persistent identifier. Authors can take an extra step by including their own CITATION file, in which the data and associated methods are described in detail. When citing data, you should cite the preferred CITATION file in the first instance, else, if there is a data publication, (for example in _Scientific Data_), cite this, and if none of those apply, cite the repository in which the data is deposited (e.g., a Zenodo repository).
